A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation

### 2016-06-21 crystal-gel/talk_Roscoff.tex
phase inversion

		\node[text width=6em, anchor=base east, inner xsep=0,xshift=1.5\mrad, align=right] (coll) at (colli.east|-pur.base) {depletion};
	\end{tikzpicture}\\

	\bigskip
	\includegraphics[width=\textwidth]{../CNRS-2015/Renth_sedim.jpg}\\
	\textit{\footnotesize Poon PRL 1999, Renth PRE 2001}
	\begin{itemize}
	\item triple coexistence
	\item various kinetic paths
	\end{itemize}

	\column{0.45\textwidth}
	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{presentation/Renth_diag.pdf}
	\end{columns}
\end{frame}

\begin{frame}{Gel formation}
	\begin{columns}
		\column{0.5\textwidth}
		\vspace{\baselineskip}
		\tikzsetnextfilename{phdiagth}%
		\begin{tikzpicture}[
			every axis/.style={
			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
			}]
		\pgfplotscreateplotcyclelist{samples}{%
		blue,mark=square*,\\%
		cyan,mark=square\\%
		red, thick, mark=*\\%
		red!65!yellow, mark=otimes\\%
		red!30!yellow, mark=oplus\\%
		red!65!black, mark=o\\%
		}%
		\begin{groupplot}[
			group style={
					group name=g, group size=1 by 2,
					horizontal sep=4em,
					vertical sep=0.5em,
					x descriptions at=edge bottom,
				},
			cycle list name=samples,
			width=\textwidth,
			%height=0.5\linewidth,
			axis on top,
		]
		\nextgroupplot[
			ylabel={attraction},
			xmin=0,xmax=0.74,
			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {crystal}  -| (rel axis cs:1,0) \closedcycle;
			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
			\addplot[black, no marks] file {res_gasliquid_bl.phd} node[pos=0.5,sloped,below] {liquid};
			\addplot[black, no marks] file {res_gasliquid_bg.phd} node[pos=0.1,sloped,below] {gas};

			\node at (axis cs:0.35,2000) {g-l spinodal};

		\nextgroupplot[
			xlabel={colloid volume fraction}, ylabel={P concent.  (\si{\milli\gram/\gram})},
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} %node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  
-| (rel axis cs:1,0) \closedcycle;
			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
			\addplot[black, no marks] file {gasliquid_bl.phd} node[pos=1,sloped,below] {liquid};
			\addplot[black, no marks] file {gasliquid_bg.phd} node[pos=0.3,sloped,below] {gas};

			%\draw[->] (x.base) --(rel axis cs:0.98,0.01);
		\end{groupplot}
		\node[anchor=north west, rotate=90, inner xsep=0,lightgray] (x) at (g c1r2.south east) {crystal};
		%\draw(g c1r1.outer south west) -- +(\textwidth,0);
		\end{tikzpicture}
		\column{0.5\textwidth}
		Small polymer/colloid size ratio
		\begin{itemize}
		\item metastable gas-liquid
		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
		\item continuous minority phase
		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
		\item due to dynamic assymetry
		\end{itemize}
		\tikzsetnextfilename{vps}%
		\begin{tikzpicture}
	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.05, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.54, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|} coordinate[pos=1] (tf) node[anchor=north east,font=\footnotesize]{time};
	\let\side\relax
	\newlength\side
	\setlength{\side}{0.275\textwidth}
	\fill[ilmcolor!80!ilmorange] ($(t1)!0.1!45:(t2)$) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmorange!80!ilmcolor] ++($(cse)!0.5!(cnw)$) 
		+(60:0.3\side) circle (0.22\side) 
		+(180:0.3\side) circle (0.22\side) 
		+(-60:0.3\side) circle (0.22\side) 
		+(0:0.6\side) circle (0.22\side)
		+(120:0.6\side) circle (0.22\side)
		+(-120:0.6\side) circle (0.22\side)
		+(45:0.8\side) circle (0.22\side)
		+(-45:0.8\side) circle (0.22\side);
	\end{scope}
	\fill[ilmcolor] (cse-|t2) ++(-0.5\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmorange] ++($(cse)!0.5!(cnw)$) 
		+(60:0.3\side) circle (0.26\side) 
		+(180:0.3\side) circle (0.26\side) 
		+(-60:0.3\side) circle (0.25\side) 
		+(0:0.6\side) circle (0.25\side)
		+(120:0.6\side) circle (0.25\side)
		+(-120:0.6\side) circle (0.25\side)
		+(45:0.8\side) circle (0.25\side)
		+(-45:0.8\side) circle (0.25\side)
		%enhance broken bond
		+(120:0.15\side) circle (0.07\side);
	\end{scope}
	\fill[ilmorange] (cse-|tf) ++(-\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmcolor] ++($(cse)!0.5!(cnw)$) circle (0.07\side)
		+(0:0.3\side) circle (0.07\side) 
		+(120:0.3\side) circle (0.07\side) 
		+(-120:0.3\side) circle (0.07\side) 
		+(-60:0.6\side) circle (0.07\side)
		+(60:0.6\side) circle (0.07\side)
		++(0:0.3\side)
		+(-60:0.3\side) circle (0.07\side)
		+(60:0.3\side) circle (0.07\side)
		+(120:0.6\side) circle (0.07\side)
		+(-120:0.6\side) circle (0.07\side)
		++(180:0.9\side)
		+(-60:0.3\side) circle (0.07\side)
		+(60:0.3\side) circle (0.07\side);
	\end{scope}
	\end{tikzpicture}

	\hfill$\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$ {\usebeamercolor[fg]{itemize item}\usebeamertemplate{itemize item}} arrest 
	\end{columns}
\end{frame}

\begin{frame}{Experimental}
How to do that in 3D at particle level \alert{without a space station}?
\begin{itemize}
\item index matched + \alert{\SI{2.3}{\micro\metre} colloids} + confocal
\item density match + $T$-control
\item preparation protocol without flow
\end{itemize}
\bigskip
\begin{columns}
	\column{0.265\textwidth}
		Colloids+polymers\\
		{\footnotesize no screening
		\begin{itemize}
			\item Dispersed
			\item Homogeneous
		\end{itemize}
		\bigskip
		\begin{beamercolorbox}[rounded=true]{block body}PMMA + PS\\
		in decalin + CHB $\epsilon_r\approx5$\\
		$\kappa^{-1}>\SI{10}{\micro\metre}$
		\end{beamercolorbox}}
	\column{0.4\textwidth}
		\tikzsetnextfilename{chemin_gelification}%
		\input{presentation/chemin_gelification.pgf}
	\column{0.335\textwidth}
		\tikzsetnextfilename{cell}%
		\input{presentation/cell.pgf}
\end{columns}
\end{frame}

>>>>

		\node[text width=6em, anchor=base east, inner xsep=0,xshift=1.5\mrad, align=right] (coll) at (colli.east|-pur.base) {depletion};
	\end{tikzpicture}
	\begin{block}{Various phases}
		\begin{itemize}
			\item triple coexistence\\
			\textit{\footnotesize Poon PRL 1999}
			\item gas-liquid metastable\\\hfill to gas-crystal
		\end{itemize}
	\end{block}
	\column{0.5\textwidth}
	\vspace{0.5\baselineskip}
	\hfill\tikzsetnextfilename{sedim_glx}%
	\begin{tikzpicture}
		\node[inner sep=0] (sedim){\includegraphics[height=17\baselineskip, clip=true, trim=0 250 0 0]{presentation/sedim_gas_liq_crystal_p.png}};
		\path[every node/.style={left}] (sedim.north west) -- (sedim.south west) node[pos=0.1] {gas} node[pos=0.45] {liquid} node[pos=0.8] {crystal} node[pos=0.15, ilmorange, font=\itshape] {low viscosity} node[pos=0.5, ilmcolor, font=\itshape] {high viscosity};
		\node[anchor=north east, font=\scriptsize, inner sep=1pt] at (sedim.south east) {Paddy Royall};
	\end{tikzpicture}
	\end{columns}
\end{frame}

\begin{frame}{Gel formation}
		\begin{columns}
		\column{0.5\textwidth}
		\vspace{\baselineskip}
		\tikzsetnextfilename{phdiagth}%
		\begin{tikzpicture}[
			every axis/.style={
			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
			}]
% 		\pgfplotscreateplotcyclelist{samples}{%
% 		blue,mark=square*,\\%
% 		cyan,mark=square\\%
% 		red, thick, mark=*\\%
% 		red!65!yellow, mark=otimes\\%
% 		red!30!yellow, mark=oplus\\%
% 		red!65!black, mark=o\\%
% 		}%
		\begin{groupplot}[
			group style={
					group name=g, group size=1 by 2,
					horizontal sep=4em,
					vertical sep=0.5em,
					x descriptions at=edge bottom,
				},
			%cycle list name=samples,
			width=\textwidth,
			%height=0.5\linewidth,
			axis on top,
		]
		\nextgroupplot[
			ylabel={attraction},
			xmin=0,xmax=0.74,
			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) coordinate (O) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {crystal}  -| (rel axis cs:1,0) \closedcycle;
			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
			\addplot[black, no marks] file {res_gasliquid_bl.phd} node[pos=0.5,sloped,below,ilmcolor] (l) {liquid};
			\addplot[black, no marks] file {res_gasliquid_bg.phd} node[pos=0.1,sloped,below,ilmorange] {gas};

			\node at (axis cs:0.35,2000) {g-l spinodal};
			\draw[<->, thick, dotted] (l) -- (l -| O) node[pos=0.5,fill] {};
			%glass
			\only<2>{\addplot[pattern=crosshatch, pattern color=ilmcolor,no marks, draw=none] coordinates {(0.56,0) (0.6,500) (0.56,2400)}  -- (rel axis cs:1,1)  \closedcycle;}

		\nextgroupplot[
			xlabel={colloid volume fraction}, ylabel={P concent.  (\si{\milli\gram/\gram})},
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} %node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  
-| (rel axis cs:1,0) \closedcycle;
			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
			\addplot[black, no marks] file {gasliquid_bl.phd} node[pos=1,sloped,below,ilmcolor] (l) {liquid};
			\addplot[black, no marks] file {gasliquid_bg.phd} node[pos=0.3,sloped,below,ilmorange] {gas};
			\draw[<->, thick, dotted] (l.north) -- (rel axis cs:0,1) node[pos=0.5,fill] {};
			%glass
			\only<2>{\addplot[pattern=crosshatch, pattern color=ilmcolor,no marks, draw=none] coordinates {(0.56,0) (0.6,0.125) (0.56,2)}  -- (rel axis cs:1,1)  \closedcycle;}

			%\draw[->] (x.base) --(rel axis cs:0.98,0.01);
		\end{groupplot}
		\node[anchor=north west, rotate=90, inner xsep=0,lightgray] (x) at (g c1r2.south east) {crystal};
		%\draw(g c1r1.outer south west) -- +(\textwidth,0);
		\end{tikzpicture}
		\column{0.5\textwidth}
		%Small polymer/colloid size ratio
		\structure{Dynamic assymetry}
		\tikzsetnextfilename{vps}%
		\begin{tikzpicture}
	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.05, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.54, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|} coordinate[pos=1] (tf) node[anchor=north east,font=\footnotesize]{time};
	\node[below, font=\scriptsize, text width=0.4\textwidth, align=center] at ($(t1)!0.5!(t2)$) {continuous \textcolor{ilmcolor}{minority} phase};
	\node[below, font=\scriptsize, text width=0.3\textwidth, align=center] at ($(t2)!0.4!(tf)$) {phase inversion};
	\let\side\relax
	\newlength\side
	\setlength{\side}{0.275\textwidth}
	\fill[ilmcolor!80!ilmorange] ($(t1)!0.1!45:(t2)$) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmorange!80!ilmcolor] ++($(cse)!0.5!(cnw)$) 
		+(60:0.3\side) circle (0.22\side) 
		+(180:0.3\side) circle (0.22\side) 
		+(-60:0.3\side) circle (0.22\side) 
		+(0:0.6\side) circle (0.22\side)
		+(120:0.6\side) circle (0.22\side)
		+(-120:0.6\side) circle (0.22\side)
		+(45:0.8\side) circle (0.22\side)
		+(-45:0.8\side) circle (0.22\side);
	\end{scope}
	\fill[ilmcolor] (cse-|t2) ++(-0.5\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmorange] ++($(cse)!0.5!(cnw)$) 
		+(60:0.3\side) circle (0.26\side) 
		+(180:0.3\side) circle (0.26\side) 
		+(-60:0.3\side) circle (0.25\side) 
		+(0:0.6\side) circle (0.25\side)
		+(120:0.6\side) circle (0.25\side)
		+(-120:0.6\side) circle (0.25\side)
		+(45:0.8\side) circle (0.25\side)
		+(-45:0.8\side) circle (0.25\side)
		%enhance broken bond
		+(120:0.15\side) circle (0.07\side);
	\end{scope}
	\fill[ilmorange] (cse-|tf) ++(-\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmcolor] ++($(cse)!0.5!(cnw)$) circle (0.07\side)
		+(0:0.3\side) circle (0.07\side) 
		+(120:0.3\side) circle (0.07\side) 
		+(-120:0.3\side) circle (0.07\side) 
		+(-60:0.6\side) circle (0.07\side)
		+(60:0.6\side) circle (0.07\side)
		++(0:0.3\side)
		+(-60:0.3\side) circle (0.07\side)
		+(60:0.3\side) circle (0.07\side)
		+(120:0.6\side) circle (0.07\side)
		+(-120:0.6\side) circle (0.07\side)
		++(180:0.9\side)
		+(-60:0.3\side) circle (0.07\side)
		+(60:0.3\side) circle (0.07\side);
	\end{scope}
	\end{tikzpicture}
	\begin{itemize}
		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
		\pause
		\item glass transition $\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$
		\item arrested phase separation
		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
	\end{itemize}	
	\end{columns}
\end{frame}

\begin{frame}{Experimental}
	\vspace{\baselineskip}
	How to observe the whole process in 3D at particle level?
	\begin{itemize}
		\item index matched + \alert{\SI{2.3}{\micro\metre} colloids} + confocal
		\item density match + $T$-control
		\item preparation protocol without flow
	\end{itemize}
	\bigskip
	\begin{columns}
		\column{0.265\textwidth}
			Colloids+polymers\\
			{\footnotesize no screening
			\begin{itemize}
				\item Dispersed
				\item Homogeneous
			\end{itemize}
			\bigskip
			\begin{beamercolorbox}[rounded=true]{block body}PMMA + PS\\
			in decalin + CHB $\epsilon_r\approx5$\\
			$\kappa^{-1}>\SI{10}{\micro\metre}$
			\end{beamercolorbox}}
		\column{0.4\textwidth}
			\tikzsetnextfilename{chemin_gelification}%
			\input{presentation/chemin_gelification.pgf}
		\column{0.335\textwidth}
			\tikzsetnextfilename{cell}%
			\input{presentation/cell.pgf}
			\vspace{-\baselineskip}
			\begin{itemize}
			\item $\sim 10^4$ particles
			\item from $t=0$
			\item to $t>10^4\tau_B$
			\end{itemize}
	\end{columns}
\end{frame}
